---
title: "Influential Observations"
format: gfm
---

[Source](https://bookdown.org/sarahwerth2024/CategoricalBook/handling-influential-observations-r.html)

```{r}
#| echo: false
options(paged.print = FALSE)
```


```{r}
#| message: false
libraries <- list(
  "tidyverse", "huxtable", "janitor"
)
invisible(lapply(libraries, library, character.only = TRUE))
```


```{r}
#| message: false
nba <- read_csv("data/nba_rookie.csv") %>% 
  clean_names() %>% 
  rename_with(~ str_remove(.x, "_percent$"),
              cols = everything())
```

# Cook's Distance

A measure of how much influence an observation has on a model. The score tells how much the coefficients were changed if the observation were removed. Higher means more influence.

The threshold for "too high" is not clear. Could be 1, could be 4/N, could be 3 * mean. R uses .5 and 1.

```{r}
model <- glm(target_5yrs ~ pts + gp + fg + x3p + ast + blk,
              data = nba, family = binomial(link = logit))
```

## Plot the distrubution oc Cook's Distance

The fourth built-in diagnostic plots. `id.n` specifies the number of points to label.

```{r werth-inf-obs-1}
plot(model, which = 4, id.n = 3)
```

## Plot Residuals vs Leverage

Leverage is a measure of how far the values of the independent variables for an observation are from values the other observations.

```{r werth-inf-obs-2}
plot(model, which = 5)
```

No values are beyond the .5 threshold, so the dashed line does not show for that cutoff.

## List the influential observations

```{r}
nba %>% 
  slice(c(315, 1254, 121)) %>% 
  select(name, pts, blk)
```

## Run regression without the observations

```{r}
model2 <- glm(target_5yrs ~ pts + gp + fg + x3p + ast + blk,
              data = nba %>% slice(-c(315, 1254, 121)),
              family = binomial(link = "logit"))
```

```{r}
models <- list(model, model2)
huxreg(models)
```

Since the differences are not large, and none were past the .5 threshold, the recommendation is to not include them.



