---
title: "Linear Probability Models"
format: gfm
---


[Source](https://bookdown.org/sarahwerth2024/CategoricalBook/linear-probability-models-r.html)

```{r}
#| echo: false
options(paged.print = FALSE)
```


```{r}
#| message: false
libraries <- list("tidyverse", "ggeffects",
               "lmtest", "sandwich")
invisible(lapply(libraries, library, character.only = TRUE))
```

# Linear Probability Model

Alternative to logistic regression or probit regression.

- uses least squares regression with a binary outcome rather than continuous
- target outcome coded 0 for non-event (not present), 1 for event (present)
- coefficients refer to probability that an outcome occurs
- probabilities are always between 0 and 1, but LPMs sometimes give > 1

$$
P(Yi = 1|Xi) = \beta_0 \ + \ \beta_1X_{i,1} \ + \ \beta_2X_{i,2} + ... \beta_kX_{i,k} + \epsilon
$$

- $P(Yi = 1|Xi)$: The probability that the outcome is present (1 = present / 0 = not present)
- That probability is predicted as a linear function (meaning the effects of each coefficient are just added together to get the outcome probability).
  - $\beta_0$ - the intercept (or probability of all coefficients are 0).
  - $\beta_k$ - the probability associated with any covariate
  - $X_k$ - the value of the covariate
  - $\epsilon$ - the error (aka the difference between the predicted probability and the actual outcome [0/1])

## Model Assumptions and considerations

- __Violation of the linearity assumption__ - LPM knowingly violates the assumption that there is a linear relationship between the outcome and the covariates. If you run a lowess line, it often looks s-shaped.
- __Violation of the homoskedasticity assumption__ - Errors in a linear probability model will by nature have systematic heteroskedasticity. You can account for this by automatically using robust standard errors.
- __Normally distributed errors__ - Errors follow a normal distribution. This assumption only matters if you have a very small sample size.
- __Errors are independent__ - There is no potential clustering within your data. Examples would be data collected from withing a set of schools, within a set of cities, from different years, and so on.
- __X is not invariant__ - The covariates that you choose can’t all be the same value! You want variation so you can test to see how the outcome changes when the x value changes.
- __X is independent of the error term__ - There are no glaring instances of confounding in your model. This assumption is about unmeasured or unobserved variables that you might reasonably expect would be associated with your covariates and outcome. You can never fully rule this violation out, but you should familiarize yourself with other studies in your area to know what major covariates you should include.
- __Errors have a mean of 0 (at every level of X)__ - If the mean is positive, your model is UNDER-estimating the outcome (more of the errors are positive). If the mean is negative, your model is OVER-estimating the outcome.

## Pros and Cons

Pros:

    The LPM is simple to run if you already know how to run a linear regression.
    It is easy to interpret as a change in the probability of the outcome ocurring.

Cons:

    It violates two of the main assumptions of normal linear regression.
    You’ll get predict probabilities that can’t exist (i.e., they are above 1 or below 0).

# Running an LPM

```{r}
titanic_df <- read.csv("data/titanic_df.csv") %>% 
  as_tibble()
```

```{r}
#| warning: false
#| message: false
titanic_df %>% 
  drop_na(port) %>% 
  ggplot(aes(x = port, y = survived)) +
  geom_point() +
  geom_smooth(method = "loess", se = F) +
  theme_minimal()
```

## Run Basic Model

```{r}
titanic_df <- titanic_df %>% 
  mutate(
    port = factor(port, 
                  labels = c("Cherbourg", "Queenstown", "Southhampton")),
    female = factor(female, labels = c("Male", "Female")),
    pclass = factor(
      pclass, 
      labels = c("1st (Upper)", "2nd (Middle)", "3rd (Lower)")),
    fare_log = if_else(fare == 0, NA, log(fare)))

basic_fit <- lm(survived ~ port, data = titanic_df)

(basic_fit_robust <- coeftest(
  basic_fit, vcov = vcovHC(basic_fit, type = "HC1")
))
```

## Run model with additional covariates.

```{r}
full_fit <- lm(survived ~ port + female + fare_log,
               data = titanic_df)
(full_fit_robust <- coeftest(
  full_fit, vcov = vcovHC(full_fit, type = "HC1")
))
```

```{r}
map(2:5,
    \(x) full_fit$coefficients[x] * 100)
```

A person bording in Queenstown had a 7.6% decreased probability of surviving compared to someone boarding in Cherbourg, not statistically significant. A person bording in Sourhhampton had a 10.1% decreased probability of surviving compared to someone boarding in Cherbourg. Female passengers had a 49.5% increased probability of survival. A 1% increase in ticket fare is associated with a 9.7% increased probability of surviving.

## Plot regression line

```{r}
ggplot(titanic_df %>% drop_na(port),
       aes(x = as.numeric(port), y = survived)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  theme_minimal()
```

## Plot probabilities in margins plot

```{r}
ggpredict(full_fit, terms = "port", 
          vcov.fun = "vcovHC", vcov.type = "HC0") %>%
  plot(connect_lines = T)
```

```{r}
ggpredict(full_fit, terms = c("port", "female"), 
          vcov.fun = "vcovHC", vcov.type = "HC0") %>%
  plot(connect_lines = T)
```

## Check Assumptions

1. Linearity is violated as expected

2. Check the violation of heteroskedasticity assumption

```{r}
plot(full_fit, which = 1)
```

3. Normally distributed errors

```{r}
plot(full_fit, which = 2)
```

This is not normal, but ok due to sample size

4. Uncorrelated errors - no natural clusters

5. X is not invariant - x variables tak more than one value

6. X is independent of error term

There are confounding variables, some may be big enough to change the results.

```{r}
full_fit_p <- lm(survived ~ port + female + pclass,
               data = titanic_df)
(full_fit_robust_p <- coeftest(
  full_fit_p, vcov = vcovHC(full_fit_p, type = "HC1")
))
```

```{r}
ggpredict(full_fit_p, terms = c("port", "female"), 
          vcov.fun = "vcovHC", vcov.type = "HC0") %>%
  plot(connect_lines = T)
```







