---
title: "Margins and Graph Design"
format: gfm
---


[Source](https://bookdown.org/sarahwerth2024/CategoricalBook/review-margins-graph-design-r.html)

```{r}
#| echo: false
options(paged.print = FALSE)
```


```{r}
#| message: false
libraries <- list(
  "tidyverse", "ggeffects", 
  "janitor", "margins"
)
invisible(lapply(libraries, library, character.only = TRUE))
```

# Question

What characteristics of passengers is associated with survival?

# Margins

Margins refers to predicted values or marginal effects. 

In regression with binary outcome, the predicted value is a predicted probability. In regression with a count outcome, the predicted value is a count. Can use `ggpredict`.

A marginal effect tells how much the outcome changes based on one unit change in the variable. It is a slope, which may not be constant. Use `margins`.

```{r}
#| message: false
titanic_df <- read_csv("data/titanic_df.csv")
```

```{r}
titanic <- titanic_df %>%
  mutate(
    port = factor(
      port,
      labels = c("Cherbourg", "Queenstown", "Southhampton")
    ),
    female = factor(female, labels = c("Male", "Female")),
    pclass = factor(pclass,
      labels = c("1st (Upper)", "2nd (Middle)", "3rd (Lower)")
    )
  )
```

```{r}
fit1 <- glm(survived ~ port + female + log_fare + parch,
            data = titanic,
            family = binomial(link = "logit"))
```

# Predicted Values

Three decisions must be made:

## What is the variable of focus?

If categorical, include in terms. If continuous, must supply values for the varible to plot.

```{r mgd-werth-1}
ggpredict(fit1, terms = "port") %>% 
  plot()
```

```{r mgd-werth-2}
ggpredict(fit1, terms = "log_fare[0, 2, 4, 6]") %>% 
  plot()
```

## Add additional variable?

More than two variables gets confusing.

```{r mgd-werth-3}
ggpredict(fit1,
          terms = c("log_fare[0, 2, 4, 6]", "parch[0, 3, 6]")) %>% 
  plot()
```

```{r mgd-werth-4}
ggpredict(fit1,
          terms = c("log_fare[0, 2, 4, 6]", "port")) %>% 
  plot()
```

```{r mgd-werth-5}
ggpredict(fit1, terms = c("port", "female")) %>% 
  plot(connect_lines = TRUE)
```

## How to treat the other X variables

- Hold at means
- Use representative values
- Run with observed values and compute the average predicted value/effect

Means is the default.

```{r}
ggpredict(fit1, terms = "port")
```

For representative values, use the `condition` argument.

```{r}
ggpredict(fit1, terms = "port",
          condition = c(female = "Female", log_fare = 2, parch = 2))
```

```{r}
predict_response(fit1, "port", margin = "empirical",
                 condition = c(female = "Female", log_fare = 2, parch = 2))
```

```{r}
predict_response(fit1, "port", margin = "marginaleffects",
                 condition = c(female = "Female", log_fare = 2, parch = 2))
```

# Marginal Effects

Categorical example

```{r}
margins(fit1, variables = "port")
```

Continuous example

```{r}
margins(fit1, variables = "log_fare")
```

Continuous with specified values

```{r}
margins(fit1, variables = "log_fare",
        at = list(log_fare = c(0, 2, 4, 6)))
```

Handling the other variables.

Marginal effects at means

```{r}
margins(fit1, variables = "port",
        at = list(
          female = "Male", log_fare = mean(fit1$data$log_fare),
          parch = mean(fit1$data$parch)
        ))
```

Marginal effects with representative values

```{r}
margins(fit1, variables = "port",
        at = list(female = "Male", log_fare = 2, parch = 2))
```

Average marginal effects

```{r}
margins(fit1, variables = "port")
```

# Plot

```{r mgd-werth-6}
#| warning: false
ggpredict(fit1, terms = c("log_fare[0, 2, 4, 6]", "female")) %>%
    plot() +
    labs(
      x = "Ticket Fare (logged)",
      y = "Predicted Probability",
      title = ""
    ) +
    scale_color_brewer(palette = "Set1",
                       labels = c("Male", "Female")) +
    theme_minimal() +
    theme(
      axis.title = element_text(size = 14),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "lightgrey"),
      legend.position = c(.2,.85), 
      legend.title = element_blank(),
      legend.direction = "horizontal"
      )
```



